Staged Abstract Interpreters

### 2019-09-09 paper/30_stageconc.tex
typo

For function applications @ap_clo@, we have two \textit{next-stage} values @fun@
and @arg@. But what we can do with these values of type @Rep@ is limited. 
In fact, @fun@ does not have an intensional application operation we can use
directly. We only know that it is a current-stage representation of functions produced by @close@,
whose next-stage form is a Scala function.
With this fact, we can generate a next-stage function application for it, also in
Scala. The method @emit_ap_clo@ produces a current-stage representation
of a next-stage function application, with necessary arguments such as the latest
store @σ@. Meanwhile, we still want the method @ap_clo@ implemented in monadic
style, so that it can be smoothly connected with other parts of the
interpreter.  To achieve this, we lift the result values and stores from
the future stage into our current-stage monad stack, with a binding @vs@.
Finally, we use @put_store@ to update the store with @vs._2@ and return the
value result @vs._1@ into the current-stage monadic value.

@@

The IR nodes are manipulated by the LMS passes and finally generated as
next-stage Scala programs. To generate code, LMS provides an @emitNode@ method for
programmers to control over what code is generated for each kind of nodes. Here,
we match @IRCompiledClo@ and @IRApClo@, and generate definitions via
@emitValDef@ for them.  For compiled closures, we pass the function @f@ with its
accompanying syntactic term and environment to the next-stage value
representation @CompiledClo@. For applications, we know that @fun@ is
an instance of @CompiledClo@ at the next stage, and it has a callable field
@f@. Hence, we generate a next-stage function application @f(arg, σ)@.  The
invocations of @quote@ are used to generate proper names for the other
next-stage values.
\begin{lstlisting}
override def emitNode(sym: Sym[Any], rhs: Def[Any]) = rhs match {
  case IRCompiledClo(f, λ, ρ) => emitValDef(sym, s"CompiledClo($\textdollar${quote(f)}, $\textdollar${quote(λ)}, $\textdollar${quote(ρ)})")
  case IRApClo(fun, arg, σ) => emitValDef(sym, s"$\textdollar$fun.f($\textdollar${quote(arg)}, $\textdollar${quote(σ)})")
  ...
}
\end{lstlisting}

>>>>

For function applications @ap_clo@, we have two \textit{next-stage} values @fun@
and @arg@. But what we can do with these values of type @Rep@ is limited. 
In fact, @fun@ does not have an intensional application operation we can use
directly. We only know that it is a current-stage representation of functions produced by @close@,
whose next-stage form is a Scala function.
With this fact, we can generate a next-stage function application for it, also
in our target language Scala. The method @emit_ap_clo@ produces a current-stage
representation of a next-stage function application, with necessary arguments
such as the latest store @σ@. Meanwhile, we still want the method @ap_clo@
implemented in monadic style, so that it can be smoothly connected with other
parts of the interpreter.  To achieve this, we lift the result values and
stores from the future stage into our current-stage monad stack, with a binding
@vs@.  Finally, we use @put_store@ to update the store with @vs._2@ and return
the value result @vs._1@ into the current-stage monadic value.

@@

The IR nodes are manipulated by the LMS passes and finally generated as
next-stage Scala programs. To generate code, LMS provides an @emitNode@ method for
programmers to control over what code is generated for each kind of nodes. Here,
we match @IRCompiledClo@ and @IRApClo@, and generate definitions via
@emitValDef@ for them.  For compiled closures, we pass the function @f@ with its
accompanying syntactic term and environment to the next-stage value
representation @CompiledClo@. For applications, we know that @fun@ is
an instance of @CompiledClo@ at the next stage, and it has a callable field
@f@. Hence, we generate a next-stage function application @f(arg, σ)@.  The
invocations of @quote@ are used to generate proper names for other
next-stage values.
\begin{lstlisting}
override def emitNode(sym: Sym[Any], rhs: Def[Any]) = rhs match {
  case IRCompiledClo(f, λ, ρ) => emitValDef(sym, s"CompiledClo($\textdollar${quote(f)}, $\textdollar${quote(λ)}, $\textdollar${quote(ρ)})")
  case IRApClo(fun, arg, σ) => emitValDef(sym, s"$\textdollar$fun.f($\textdollar${quote(arg)}, $\textdollar${quote(σ)})")
  ...
}
\end{lstlisting}

### 2019-09-09 paper/40_absinterp.tex
typo

After seeing the unstaged and staged concrete interpreter, now we turn our
focus to abstract interpreters under the same framework. We first present a
lattice representation with binding-time abstraction, as well as simple
abstract domains, such as power sets.  Then, we show the abstract
components, namely, abstract values, abstract environments, and abstract stores.
The abstract interpreter we construct in this section is similar to
\citet{DBLP:journals/pacmpl/DaraisLNH17}'s. For simplicity, it is
context/path/flow-insensitive by our choice. In Section \ref{cfa}, we will
instantiate it to a commonly-used context-sensitive analysis;
\citet{Darais:2015:GTM:2814270.2814308} show how to achieve path- and
flow-sensitivity by varying the monads, which is also applicable to our
unstaged or staged abstract interpreter.

@@

In this section, we follow the abstracting definitional interpreters
(ADI) approach \cite{DBLP:journals/pacmpl/DaraisLNH17} to refactor our monadic
concrete interpreter to an monadic abstract interpreter.

@@

It is important that the abstract stores now map addresses to sets of abstract
values, indicating that an address may point to an over-approximated set of
runtime values.  This can be justified by the approximation nature and
nondeterminism during the analysis.  For example, when analyzing a
conditional expression, we may not have enough information to decide which
branch will be taken, thus a sound treatment is to explore the both branches.
Also, at some later point, we need to compute the join of two path.  Another
nondeterminism comes from function applications, for instance, @f(a)@, the
possible $\lambda$ terms of @f@ is computed approximately.  As a
result, we need to store all the possible callees, as well as we will
retrieve a set of closures when dereferencing an address.

>>>>

After seeing the unstaged and staged concrete interpreter, now we turn our
focus to abstract interpreters under the same framework. We first present a
lattice representation with binding-time abstraction, as well as simple
abstract domains, such as power sets.  Then, we show the abstract
components, namely, abstract values, abstract environments, and abstract stores.
The abstract interpreter we construct in this section is similar to
\citet{DBLP:journals/pacmpl/DaraisLNH17}'s. For simplicity, it is
context/path/flow-insensitive by our choice. In Section \ref{cfa}, we will
instantiate it to a commonly-used context-sensitive analysis;
\citet{Darais:2015:GTM:2814270.2814308} showed how to achieve path- and
flow-sensitivity by varying the monads, which is also applicable to our
unstaged or staged abstract interpreter.

@@

In this section, we follow the abstracting definitional interpreters
(ADI) approach \cite{DBLP:journals/pacmpl/DaraisLNH17} to refactor our monadic
concrete interpreter to a monadic abstract interpreter.

@@

It is important that the abstract stores now map addresses to sets of abstract
values, indicating that an address may point to an over-approximated set of
runtime values.  This can be justified by the approximation nature and
nondeterminism during the analysis.  For example, when analyzing a
conditional expression, we may not have enough information to decide which
branch will be taken, thus a sound treatment is to explore the both branches.
Also, at some later point, we need to compute the join of two path.  Another
nondeterminism comes from function applications, for instance, @f(a)@, the
possible $\lambda$ terms of @f@ are computed approximately.  As the
result, we need to store all the possible callees, as well as we will
retrieve a set of closures when dereferencing an address.
